---
title: "Amenities from Geodatabase Nacional"
date: "Last modified: `r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: show # class.source = "fold-hide" to hide code and add a button to show it
    df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
    fig_caption: true
    code_download: true
    css: visual.css
always_allow_html: true
bibliography: 
  - references.bib
  - grateful-refs.bib
---

```{r}
# Create a clipboard button on the rendered HTML page
source(here::here("clipboard.R")); clipboard
# Set seed for reproducibility
set.seed(1982) 
# Set global options for all code chunks
knitr::opts_chunk$set(
  # Disable messages printed by R code chunks
  message = FALSE,    
  # Disable warnings printed by R code chunks
  warning = FALSE,    
  # Show R code within code chunks in output
  echo = TRUE,        
  # Include both R code and its results in output
  include = TRUE,     
  # Evaluate R code chunks
  eval = TRUE,       
  # Enable caching of R code chunks for faster rendering
  cache = FALSE,      
  # Align figures in the center of the output
  fig.align = "center",
  # Enable retina display for high-resolution figures
  retina = 2,
  # Show errors in the output instead of stopping rendering
  error = TRUE,
  # Do not collapse code and output into a single block
  collapse = FALSE
)
# Start the figure counter
fig_count <- 0
# Define the captioner function
captioner <- function(caption) {
  fig_count <<- fig_count + 1
  paste0("Figure ", fig_count, ": ", caption)
}
```


```{r}
library(sf)
library(tmap)
library(mapview)
library(dplyr)
# set mapview options
mapviewOptions(basemaps = c("CartoDB.Positron",
                            "OpenStreetMap",
                            "Esri.WorldImagery",
                            "OpenTopoMap"))
```

# Introduction

In this section, we will explore the amenities available in the Geodatabase Nacional of Ecuador. The data was obtained from this [website](https://www.ecuadorencifras.gob.ec/documentos/web-inec/Geografia_Estadistica/Micrositio_geoportal/index.html). On the page, go to the bottom section (called Cartografia Historica) of the page and click on `2022` and then on the `Descarga de Marco Geoestadístico 2021` button. This will download a file named `GEODATABASE_NACIONAL_2021.zip`. Unzip the file and you will find a folder named `GEODATABASE_NACIONAL_2021`. Inside this folder, there is a file named `GEODATABASE_NACIONAL_2021.gpkg`, which is a geopackage file containing several (9) layers. Using the [`manabi_area_simple.RDS`](https://github.com/leninrafaelrierasegura/manabi/blob/main/clean_data/manabi_area_simple.RDS) file, I filtered the contents of each layer.

---

By the way, the number of the layers are:

```{r}
LAYERS_name <- c("viv_p", "ejes_l", "zon_a", "sec_a", "loc_p", "ingresos_l", "ca04_a", "aream_a", "man_a")
```

---

Most of the layers do not contain useful information for our purposes. Here are some details about the layers:

- `aream_a`: This layer seems to contain the head of the administrative areas as polygons, such as provinces, cantons, and parishes.
- `loc_p`: This layer contains the locations of small towns and villages, apparently.
- `ca04_a`: This layer contains the locations of amenities, such as schools, hospitals, and other public buildings.


## Plot the area of interest

```{r}
area_of_interest <- readRDS("clean_data/manabi_area_simple.RDS")
mapview(
  area_of_interest,
  zcol = "geometry",        # attribute used for fill
  alpha.regions = 0,    # fill transparency
  color = "black",        # border color
  alpha = 0.4,            # border transparency
  legend = FALSE           # remove legend
)
```

# layer `aream_a`



```{r, out.width = "100%", fig.cap = captioner("Area of the head of the administrative units.")}
aream_a <- readRDS("clean_data/aream_a_cropped_to_manabi.RDS")
mapview(
  aream_a,
  zcol = "tipo_aream",        # attribute used for fill
  alpha.regions = 0.4,    # fill transparency
  color = "black",        # border color
  alpha = 1,            # border transparency
  legend = TRUE           # remove legend
)
```

# layer `loc_p`


```{r, out.width = "100%", fig.cap = captioner("Small towns and villages, apparently.")}
loc_p <- readRDS("clean_data/loc_p_cropped_to_manabi.RDS")
mapview(
  loc_p,
  zcol = "n_loc",        # attribute used for fill
  alpha.regions = 0.4,    # fill transparency
  color = "black",        # border color
  alpha = 0.2,            # border transparency
  cex = 1.5, #           points size
  
  legend = FALSE           # remove legend
) 
```


# layer `ca04_a`

```{r}
# ca04_a <- readRDS("clean_data/ca04_a_cropped_to_manabi.RDS")
# amn <- unique(ca04_a$cod_otros) # check the unique values of the cod_otros column
# amn
#  [1] NA                         "EDIFICIO EDUCACIONAL"     "TEMPLO RELIGIOSO"        
#  [4] "CASA COMUNAL"             "EDIFICIO IMPORTANTE"      "EDIFICIO DE REFERENCIA"  
#  [7] "CAMPO DEPORTIVO"          "CEMENTERIO"               "GASOLINERA"              
# [10] "ESTABLECIMIENTO DE SALUD" "PARQUE"                   "PLAZA"                   
# [13] "TANQUE DE AGUA"           "EDIFICIO REFERENCIA" 
# for (i in 2:length(amn)) {
#   aux <- filter(ca04_a, cod_otros == amn[i])
#   saveRDS(aux, paste0("clean_data/amn_", gsub(" ", "_", amn[i]), ".RDS"))
# }
```

Here we can see the `EDIFICIO EDUCACIONAL` amenities.

```{r, out.width = "100%", fig.cap = captioner("Educational buildings.")}
amn_EDIFICIO_EDUCACIONAL <- readRDS("clean_data/amn_EDIFICIO_EDUCACIONAL.RDS")
mapview(
    amn_EDIFICIO_EDUCACIONAL,
    zcol = "nom_edif",        # attribute used for fill
    alpha.regions = 0.4,    # fill transparency
    color = "black",        # border color
    alpha = 1,            # border transparency
    legend = FALSE           # remove legend
  ) 
```




# References 

```{r}
grateful::cite_packages(output = "paragraph", out.dir = ".")
```




