---
title: "Elevation"
date: "Last modified: `r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: show # class.source = "fold-hide" to hide code and add a button to show it
    df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
    fig_caption: true
    code_download: true
    css: visual.css
always_allow_html: true
bibliography: 
  - references.bib
  - grateful-refs.bib
---

```{r}
# Create a clipboard button on the rendered HTML page
source(here::here("clipboard.R")); clipboard
# Set seed for reproducibility
set.seed(1982) 
# Set global options for all code chunks
knitr::opts_chunk$set(
  # Disable messages printed by R code chunks
  message = FALSE,    
  # Disable warnings printed by R code chunks
  warning = FALSE,    
  # Show R code within code chunks in output
  echo = TRUE,        
  # Include both R code and its results in output
  include = TRUE,     
  # Evaluate R code chunks
  eval = TRUE,       
  # Enable caching of R code chunks for faster rendering
  cache = FALSE,      
  # Align figures in the center of the output
  fig.align = "center",
  # Enable retina display for high-resolution figures
  retina = 2,
  # Show errors in the output instead of stopping rendering
  error = TRUE,
  # Do not collapse code and output into a single block
  collapse = FALSE
)
# Start the figure counter
fig_count <- 0
# Define the captioner function
captioner <- function(caption) {
  fig_count <<- fig_count + 1
  paste0("Figure ", fig_count, ": ", caption)
}
```


```{r}
library(sf)
library(mapview)
library(dplyr)
library(terra)
# set mapview options
mapviewOptions(basemaps = c("CartoDB.Positron",
                            "OpenStreetMap",
                            "Esri.WorldImagery",
                            "OpenTopoMap"))
```



# Introduction

Data was downloaded from [OpenTopography](https://portal.opentopography.org/raster?opentopoID=OTSDEM.032021.4326.3). There are two resolutions available, 30m and 90m. You get a GeoTIFF file for each resolution.


Go to this [file](https://github.com/leninrafaelrierasegura/manabi/blob/main/elevation.R)  to see how the data was handled.

Here is a summary of the steps taken:

- The data was clipped to the study area using the polygon in [`manabi_area_simple.RDS`](https://github.com/leninrafaelrierasegura/manabi/blob/main/clean_data/manabi_area_simple.RDS). Two files are produced:  [`dem30_manabi.tif`](https://github.com/leninrafaelrierasegura/manabi/blob/main/clean_data/dem30_manabi.tif) and [`dem90_manabi.tif`](https://github.com/leninrafaelrierasegura/manabi/blob/main/clean_data/dem90_manabi.tif).

- The resulting dataset was classified into elevation ranges: 0-50m, 50-150m, 150-250m, 250-350m, 350-450m, 450-750m, and above 750m. The resulting regions are stored in [`dem_polygons_manabi.RDS`](https://github.com/leninrafaelrierasegura/manabi/blob/main/clean_data/dem_polygons_manabi.RDS).


# Elevation data

```{r, out.width = "100%"}
dem90_mask <- rast("clean_data/dem90_manabi.tif")
mapview(dem90_mask, col.regions = terrain.colors(100), maxpixels = 4e+05)
```

# Elevation ranges

```{r, out.width = "100%"}
elevation_ranges <- readRDS("clean_data/dem_polygons_manabi.RDS")
mapview(dplyr::filter(elevation_ranges, elevation_range != "0: NoData/Below 0"), zcol = "elevation_range", alpha.regions = 0.5)
```

150-250m is the sweet spot.

Known spots for reference:

- [El Rosario](https://mapcarta.com/es/19661918) ~ 200m
- [La polvareda](https://mapcarta.com/es/W435020166) ~ 200m
- El Congoma ~ 200m
- [La Montufar](https://mapcarta.com/es/N2506430045) ~ 640m
- [Las Mercedes](https://mapcarta.com/es/29290582) ~ 750m


# References 

```{r}
grateful::cite_packages(output = "paragraph", out.dir = ".")
```




