---
title: "Temperature"
date: "Last modified: `r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: show # class.source = "fold-hide" to hide code and add a button to show it
    df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
    fig_caption: true
    code_download: true
    css: visual.css
always_allow_html: true
bibliography: 
  - references.bib
  - grateful-refs.bib
---

```{r}
# Create a clipboard button on the rendered HTML page
source(here::here("clipboard.R")); clipboard
# Set seed for reproducibility
set.seed(1982) 
# Set global options for all code chunks
knitr::opts_chunk$set(
  # Disable messages printed by R code chunks
  message = FALSE,    
  # Disable warnings printed by R code chunks
  warning = FALSE,    
  # Show R code within code chunks in output
  echo = TRUE,        
  # Include both R code and its results in output
  include = TRUE,     
  # Evaluate R code chunks
  eval = TRUE,       
  # Enable caching of R code chunks for faster rendering
  cache = FALSE,      
  # Align figures in the center of the output
  fig.align = "center",
  # Enable retina display for high-resolution figures
  retina = 2,
  # Show errors in the output instead of stopping rendering
  error = TRUE,
  # Do not collapse code and output into a single block
  collapse = FALSE
)
# Start the figure counter
fig_count <- 0
# Define the captioner function
captioner <- function(caption) {
  fig_count <<- fig_count + 1
  paste0("Figure ", fig_count, ": ", caption)
}
```


```{r}
library(sf)
library(mapview)
library(dplyr)
library(terra)
# set mapview options
mapviewOptions(basemaps = c("CartoDB.Positron",
                            "OpenStreetMap",
                            "Esri.WorldImagery",
                            "OpenTopoMap"))
```



# Introduction



Data was downloaded from [WorldClim](https://www.worldclim.org/) using the `geodata` package. The data represents average monthly temperature at a spatial resolution of 30 arc-seconds (approximately 1 km² at the equator). You get 12 raster layers, one for each month of the year. The data is in degrees Celsius and it was collected from 1970 to 2000.

Go to this [file](https://github.com/leninrafaelrierasegura/manabi/blob/main/temperature.R)  to see how the data was handled.


Here is a summary of the steps taken:

- For the monthly average temperature data, we calculated the mean across the 12 layers to get the average climatology.

- For the monthly minimum temperature data, we calculated the minimum across the 12 layers.

- For the monthly maximum temperature data, we calculated the maximum across the 12 layers.

- We then clipped the rasters to the area of interest (Manabí province) using a polygon shapefile.

- Finally, we applied a condition to identify suitable areas where the minimum temperature is greater than or equal to 18°C and the maximum temperature is less than or equal to 30°C. (this is done in this file)


```{r}
manabi_avg_temp <- rast("clean_data/manabi_avg_temp.tif")
manabi_min_temp <- rast("clean_data/manabi_min_temp.tif")
manabi_max_temp <- rast("clean_data/manabi_max_temp.tif")

# Compute global range
all_min <- min(minmax(manabi_avg_temp), minmax(manabi_min_temp), minmax(manabi_max_temp))
all_max <- max(minmax(manabi_avg_temp), minmax(manabi_min_temp), minmax(manabi_max_temp))

all_min; all_max
```

# Avearage Temperature

```{r, out.width = "100%", fig.height = 12}
mapview(manabi_avg_temp, layer.name = "Avg Temp (°C)", at = seq(all_min, all_max, length.out = 101), legend = FALSE)
```


# Minimum Temperature

```{r, out.width = "100%", fig.height = 12}
mapview(manabi_min_temp, layer.name = "Min Temp (°C)", at = seq(all_min, all_max, length.out = 101), legend = FALSE)
```

# Maximum Temperature

```{r, out.width = "100%", fig.height = 12}
mapview(manabi_max_temp, layer.name = "Max Temp (°C)", at = seq(all_min, all_max, length.out = 101), legend = FALSE)
```

# Suitable area

```{r}
suitable_area <- (manabi_min_temp >= 19.5) & (manabi_max_temp <= 30)

suitable_min <- mask(manabi_min_temp, suitable_area, maskvalues = 0)
suitable_max <- mask(manabi_max_temp, suitable_area, maskvalues = 0)

# Stack the two rasters into a single multi-layer raster
suitable_stack <- c(suitable_min, suitable_max)

# Optional: rename the layers for clarity
names(suitable_stack) <- c("Tmin", "Tmax")

```

Below we show the minimum temperature layer of the suitable area. The important thing is the area.


```{r, out.width = "100%", fig.height = 12}
mapview(suitable_stack, layer.name = "Suitable Min Temp (°C)", legend = TRUE)
```



# References 

```{r}
grateful::cite_packages(output = "paragraph", out.dir = ".")
```



